FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions

### 2023-08-09 paper.tex
Lots more work in related work section

Error analysis, typical answer.
Difficult.
CITE

Toronto and McCarthy\cite{torontoPracticallyAccurateFloatingPoint2014} developed a methodology for assessing where arithmetic operations are liable to magnify \fp{} error.
They do not provide any tooling to find \emph{where} the error is causing problems, but they do provide guidance on what to do to mitigate error.

Herbie\cite{panchekhaAutomaticallyImprovingAccuracy2015} is a tool for automatically rewriting arithmetic expressions to reduce \fp{} error.
\FT{} tracks down exceptional \fp{} values; once the location of a gen is found, Herbie could be employed to rewrite the arithmetic expression, if \fp{} error is the cause of the exceptional value.
Odyssey\cite{misbackOdysseyInteractiveWorkbench2023} provides a helpful interactive interface to the tooling that Herbie provides.

% Herbie, improving accuracy automatically.
% ParetoHerbie

%% http://eprints.maths.manchester.ac.uk/2873/1/fami22.pdf
%% https://drive.google.com/drive/folders/1HkgMpxi6hsqLSZj9tSnuCxnFExPPjmvx?usp=sharing
%% https://docs.google.com/presentation/d/1ueqzb5QY9YHmtY2DfY_hpN9oufSOcXErLWsJOZQxgzo/edit?usp=sharing
%% https://ieeexplore.ieee.org/document/8916392
%% https://link.springer.com/chapter/10.1007/978-3-319-73721-8_24

Demmel et al.\cite{ddghlllprr-correctness-2022} examine \fp{} exceptional value handling in the BLAS and LAPACK libraries and propose solutions to inconsistencies.
They also suggest a fuzzing technique that we build upon with \FT{}:
instead of fuzzing the \emph{inputs} to a program, we fuzz by randomly injecting exceptional values at \emph{any} point where there is an arithmetic operation. Demmel suggests that this approach is useful since any \fp{} operation could lead to an exceptional value if the inputs are in dangerous ranges.

\texttt{Sherlogs}\footnote{\url{https://github.com/milankl/Sherlogs.jl}} by Milan Klöwer inspired the use of a custom number type to intercept operations on a number.
In contrast to \FT{} which monitors for exceptional values and logs stack traces at interesting points in their lifetime, \texttt{Sherlogs} tracks and reports the range of values seen over the course of a computation.
This is intended to provide insight into whether or not a library could tolerate a lower-precision \fp{} format.

FPSpy x86 binaries [10]
Dinda etal hpdc 2020

FPChecker, " considers CPU OpenMP and MPI codes, again using LLVM
instrumentation"~\cite{ltlg-iiswc-2022}
% FPChecker has a nice table of tools; some already mentioned here

6-8,24,25
Arnab Das scalable yet rigorous
Marc Daumas Certification of Bounds Expres
David Delmas Towards an Industrial Use of
Alexey Solovyev FPTaylor Results table
Titolo absint

CSTG~\cite{humphreySystematicDebuggingMethods2014}.
What else???

GPU-FPX state of art, companion project~\cite{llsflg-hpdc-2023}.
Latest in line of work.
BinFPE exn checking, slower than GPU-FPX, 
\cite{llg-soap-2022}.
FPChecker, early pioneer, fpx in gpus, llvm-level instrumentation
if compiled with Clang, but most gpus use closed NVCC compiler~\cite{l-ase-2019}.

@@

% Local Variables:
% jinx-local-words: "CFL CSTG's CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq RxInfer ShallowWaters Sherlogs TrackedFloat isnan"
% citar-bibliography: ("./ref.bib")
% End:

>>>>

\subsection{Floating-point error analysis}

Demmel et al.\cite{ddghlllprr-correctness-2022} examine \fp{} exceptional value handling in the BLAS and LAPACK libraries and propose solutions to inconsistencies.
Their solution includes a characterization equivalent to our gen-prop-kill framework. % pg. 18---good way to cite this page specifically?
They also suggest a fuzzing technique that we realize in \FT{}:
instead of fuzzing the \emph{inputs} to a program, we fuzz by randomly injecting exceptional values at \emph{any} point where there is an arithmetic operation. Demmel suggests that this approach is useful since any \fp{} operation could lead to an exceptional value if the inputs are in dangerous ranges.

Toronto and McCarthy\cite{torontoPracticallyAccurateFloatingPoint2014} developed a methodology for assessing where arithmetic operations are liable to magnify \fp{} error.
Their work includes error visualization and work on how to avoid error-magnifying ``badlands''.
They focus on rewriting arithmetic expressions to avoid error which could lead to exceptional values, rather than on exceptional values themselves.

In a similar vein, Herbie\cite{panchekhaAutomaticallyImprovingAccuracy2015} is a tool for automatically rewriting arithmetic expressions to reduce \fp{} error.
\FT{} tracks down exceptional \fp{} values; once the location of a gen is found, Herbie could be employed to rewrite the arithmetic expression, if \fp{} error is the cause of the exceptional value.
Odyssey\cite{misbackOdysseyInteractiveWorkbench2023} provides a helpful interactive interface to the tooling that Herbie provides.

\subsection{Diagnosing floating-point exceptions}

The Julia library \texttt{Sherlogs}\footnote{\url{https://github.com/milankl/Sherlogs.jl}} by Milan Klöwer inspired the use of a custom number type to intercept operations on a number.
In contrast to \FT{} which monitors for exceptional values and logs stack traces at interesting points in their lifetime, \texttt{Sherlogs} tracks and reports the range of values seen over the course of a computation.
This is intended to provide insight into whether or not a library could tolerate a lower-precision \fp{} format.

% TODO: make sure that what I said about what FPSpy can track is true.
FPSpy~\cite{dindaSpyingFloatingPoint2020} is an \texttt{LD_PRELOAD} shared library that works on unmodified x86 binaries.
It can monitor a program during execution for any \fp{} arithmetic issues, such as division by zero, denormalization, underflow, and overflow.
\FT{} in contrast focuses on the lifetime of exceptional values; \FT{} can notice when a \NaN{} gets killed, for instance, while that is not an event that FPSpy appears to be able to track.
FPSpy has the advantage of being lightweight enough to run on production code for certain loads.

\subsection{Stack trace graphs}

CSTG~\cite{humphreySystematicDebuggingMethods2014} is built on the work of and related to the STAT~\cite{arnoldStackTraceAnalysis2007} tool developed by Laurence Livermore National Laboratory.
Please see the CSTG paper for more on the differences between the two.

\subsection{Floating-point exceptions on the GPU}

\GPUFPX{}…

% Error analysis, typical answer.
% Difficult.
% CITE

% % Herbie, improving accuracy automatically.
% % ParetoHerbie

% %% http://eprints.maths.manchester.ac.uk/2873/1/fami22.pdf
% %% https://drive.google.com/drive/folders/1HkgMpxi6hsqLSZj9tSnuCxnFExPPjmvx?usp=sharing
% %% https://docs.google.com/presentation/d/1ueqzb5QY9YHmtY2DfY_hpN9oufSOcXErLWsJOZQxgzo/edit?usp=sharing
% %% https://ieeexplore.ieee.org/document/8916392
% %% https://link.springer.com/chapter/10.1007/978-3-319-73721-8_24

% FPChecker, " considers CPU OpenMP and MPI codes, again using LLVM
% instrumentation"~\cite{ltlg-iiswc-2022}
% % FPChecker has a nice table of tools; some already mentioned here

% 6-8,24,25
% Arnab Das scalable yet rigorous
% Marc Daumas Certification of Bounds Expres
% David Delmas Towards an Industrial Use of
% Alexey Solovyev FPTaylor Results table
% Titolo absint

% GPU-FPX state of art, companion project~\cite{llsflg-hpdc-2023}.
% Latest in line of work.
% BinFPE exn checking, slower than GPU-FPX, 
% \cite{llg-soap-2022}.
% FPChecker, early pioneer, fpx in gpus, llvm-level instrumentation
% if compiled with Clang, but most gpus use closed NVCC compiler~\cite{l-ase-2019}.

@@

% Local Variables:
% jinx-local-words: "CFL CSTG's CSTGs Courant Demmel Friedrichs Herbie JuliaHub Klöwer Lewy OrdinaryDiffEq RxInfer ShallowWaters Sherlogs TrackedFloat isnan"
% citar-bibliography: ("./ref.bib")
% End:
